% Chapter Template

\chapter{Conclusion and Future Work} % Main chapter title

\label{Chapter7} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{Chapter7}

\lhead{Chapter 7. \emph{Conclusion and Future Work}} % Change X to a consecutive number; this is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Conclusion}

Conclusion: This project delivers a full‑stack Hospital Management System that unifies patient, doctor, staff, and admin workflows. Core contributions include secure authentication, unified appointment management (in‑person and virtual), EMR-linked operational modules (lab, pharmacy, inpatient, home visits), and an analytics dashboard for revenue, visits, and staff performance. The implementation emphasizes security, usability, and maintainability with a modern TypeScript codebase and well‑structured REST APIs backed by Supabase.

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------
\subsection{Outcomes and Impact}
Outcomes and impact: The HMS reduces administrative overhead through centralized scheduling and billing, improves communication with chat and video, and equips administrators with actionable insights. Reusable components and clear service boundaries accelerate future development and institutional adoption.

%-----------------------------------
%	SUBSECTION 2
%-----------------------------------

\subsection{Limitations}
Limitations: Load testing and advanced observability are out of scope; interoperability is limited to exports rather than live HL7/FHIR exchange; and predictive analytics are not yet implemented. These constraints inform the roadmap for future enhancements.

%----------------------------------------------------------------------------------------
%	SECTION 2
%----------------------------------------------------------------------------------------

\section{Future Work}

Future work: Key extensions include HL7/FHIR-based interoperability, advanced EMR features (order sets, allergies, medication reconciliation), automated reminders and escalation policies, comprehensive load testing and SLOs, mobile apps for patients and clinicians, and machine‑learning‑assisted analytics for demand forecasting and triage support.